Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index and the Use of Machine Learning Methods to Forecast the Price of the FTSE 250 Index

### 2024-08-22 tex/paper.tex
Introduction revised

The stock market has played a central role in the economic and 
political ascent of many nations ever since its inception (\citeboth{levine2002}).  
However, whenever stock market instabilities have arisen in history they 
have had far-reaching destabilising effects. The Mississippi Bubble was 
an early example of the inherent dangers within stock markets in the form of 
speculative bubbles (\citeboth{mississippibubble}). The role of
the macroeconomy in this crash was central. The French central bank, the 
\textit{Banque Royale}, aimed to convert its vast national debt into shares through the issuance of large 
amounts of government bonds and paper notes. This considerably increased the money supply, in turn 
greatly facilitating the ease of obtaining credit by individuals, many of whom invested in the Mississippi Company.
A mass sell-off began after the \textit{Banque Royale} announced the number of paper notes exceeded the 
value of metal coinage it possessed. This large-scale expansion of the money supply led to 
inflation burgeoning, reaching $23\%$ at its peak, and devastating the French economy (\citeboth{moen2001john}).  Nonetheless, stock markets also provide essential benefits by allowing companies to 
raise capital, offering investment opportunities for individuals, and 
supporting economic growth through efficient capital allocation. 
They enhance corporate governance through transparency and accountability, 
facilitate price discovery and market liquidity, and contribute to government 
financing through taxes (\citeboth{shahbaz2016}). Effective and sustainable
macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
for economic policymakers to understand the impact of macroeconomic changes 
on the stock market.

The sensitivity of the stock market to macroeconomic indicators is also of great 
relevance to investors. Macroeconomic indicators provide 
insights into the overall health and direction of the 
economy, which affect investor sentiment and in turn influence market trends.
Changes in the macroeconomy have an effect on both consumers and firms 
which impacts their profitability, because these signals 
help investors anticipate future economic conditions and corporate profits
and consequently investment returns. Positive indicators tend to boost stock prices, while 
negative indicators can lead to declines in stock prices.

@@

A key task which economic policy makers and investors undertake is to 
understand and predict how economic changes influence stock market behaviour, 
which is critical for making informed policy and investment decisions. 
This modelling of the impact of macroeconomic indicators on stock markets is typically achieved by a variety of methods: statistical analysis, econometric modelling, scenario planning, and advanced machine learning techniques. However, it is fraught with challenges that require careful consideration, advanced techniques, and a deep understanding of both economics and financial markets. In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics. 
In addition, certain macroeconomic indicator data is often so small in size that it precludes the use of certain deep learning neural network methods for prediction purposes.

@@

The remainder of the paper is organised as follows. Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables. Section 3 covers the data utilised in this study. Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved. Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE250 based on the selected macroeconomic variables. Section 6 concludes the study.

>>>>

\subsection{Context}

The stock market has played a central role in the economic and 
political ascent of many nations ever since its inception (\citeboth{levine2002}).  
However, whenever stock market instabilities have arisen in history they 
have had far-reaching destabilising effects. The Mississippi Bubble was 
an early example of the inherent dangers within stock markets in the form of 
speculative bubbles (\citeboth{mississippibubble}). The role of 
macroeconomic policy in this crash was central, and the effects of the crash 
devastated the French economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
raise capital, offering investment opportunities for individuals, and 
supporting economic growth through efficient capital allocation. 
They enhance corporate governance through transparency and accountability, 
facilitate price discovery and market liquidity, and contribute to government 
financing through taxes (\citeboth{shahbaz2016}). Effective and sustainable
macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
for economic policymakers to understand the impact of macroeconomic changes 
on the stock market.

The sensitivity of the stock market to macroeconomic indicators is also of great 
relevance to investors. Macroeconomic indicators provide 
insights into the overall health and direction of the 
economy, which affect investor sentiment and in turn influence market trends.
Changes in the macroeconomy have an effect on both consumers and firms 
which impacts their profitability, because these signals 
help investors anticipate future economic conditions and corporate profits
and consequently investment returns.

@@

Modelling the impact of macroeconomic indicators on stock markets 
typically involves a variety of methods: statistical analysis, 
econometric modelling, scenario planning, and advanced machine learning 
techniques. However, this process contains challenges that require careful 
consideration. In particular, there is a challenge to 
find a subset of macroeconomic indicators which are most relevant in 
stock market dynamics. Additionally, there is the task of
choosing the most appropriate model based on the 
research objectives and data to ensure reliable
and accurate results.

\subsection{Research objectives}

@@

\subsection{Organisation of paper}
The remainder of the paper is organised as follows. Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables. Section 3 covers the data utilised in this study. Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved. Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE250 based on the selected macroeconomic variables. Section 6 concludes the study.
